### 2024-05-03 notes/For the Group.Rmd
Convert code from Julia to RMarkdown. Finished cont, cont, fix and cont, bin, fix

---
title: "Exact Asymptotics for Causal Mediation Analysis"
author: "William Ruth"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
# Force text wrapping for long lines
library(formatR)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

library(kableExtra)

knitr::opts_chunk$set(echo = TRUE)

```

I'm doing some Monte Carlo to verify the new SE formulas. Recall that we're doing mediation analysis, so we've got a response, $Y$, an exposure, $X$, and a mediation, $M$. We also have some number of confounders, which will be grouped together in the matrix $W$. Broadly speaking, we fit two regression models, one to predict $M$ using $X$ and $W$, the other to predict $Y$ using $M$, $X$ and $W$. We then compute the mediation effect (specifically, the total effect of $X$ on $Y$) as a function of the coefficients from these two regression models. An asymptotic SE for our mediation effect estimator can then be obtained from the asymptotic standard errors of our fitted regression coefficients using the $\delta$-method.

So far, so simple. There are a few places that things start to get more complicated. First, each of the two regression models can be either linear of logistic depending on whether the corresponding response variable is continuous or binary\footnote{In principle, we could have $Y$ and/or $M$ follow any distribution with a suitable GLM formulation. I don't think I've ever seen count data (i.e. Poisson regression) used here, much less anything more exotic.}. Furthermore, we can add random effects to our regression models. In the trust study, we have random effects for the intercept, $X$ and $M$ (naturally, the latter only applies when predicting $Y$).

In each setting, I generate $X$ and three confounders, $W = [W_1, W_2, W_3]$, each iid $\mathrm{N}(0,1)$. In both models, I set regression slopes to 1, and choose the intercept so that the mean of the linear predictor is approximately zero\footnote{In the fully continuous case, the mean linear predictor is exactly zero. When the mediator is binary, I choose the $Y$-intecept as though $\mathbb{E}M=\mathrm{expit}(0)=0.5$}. I generate some datasets (typically 1000, can be adjusted as necessary) with each of $n=100, 1000$ and $10000$ observations. On each dataset, we estimate the mediation effect and its SE (using the $\delta$-method). We then compare the empirical SE (i.e. SD over the $M$ Monte Carlo replicates) to the mean and median of our $M$ estimated SEs.

In summary, we can setup each model with the same boilerplate code as follows:

```{r, boilerplate, eval=FALSE}
# Note: This code chunk is not run here, but it is called and run at the start of every subsequent section. Any changes made here will be duplicated throughout the rest of the document.

set.seed(1)

num_reps = 1000

all_Ns = c(100, 200, 500, 1000, 2000, 5000, 10000)
p_conf = 3  # Number of confounders

# Regression coefficients
## Intercepts are set separately for each model so that the mean of the linear predictor is approximately zero

## M model
a_1 = 1                       # Coefficient for X
A_2 = rep(1, times = p_conf)  # Coefficients for confounders

# Y model
b_1 = 1                       # Coefficient for M
b_2 = 1                       # Coefficient for X
B_3 = rep(1, times = p_conf)  # Coefficients for confounders

# Containers for output
all_a_hats = list()
all_b_hats = list()

all_a_SEs = list()
all_b_SEs = list()

all_med_hats = list()
all_med_SEs = list()

```

In the rest of this document, we carry our Monte Carlo studies for different model configurations. We start with the simplest case: continuous response, continuous mediator, and fixed-effects. We then move to a binary mediator, followed by binary response. Finally, we introduce random effects.

# Continuous Response, Continuous Mediator, Fixed-Effects

```{r, include=FALSE}
# Run the boilerplate setup
<<boilerplate>>
```

## Generate data

Continuous variables are modelled as linear predictor plus residual. We therefore need to set the residual variance.

```{r}
sigma_M = 0.2
sigma_Y = 0.2
```

We also need to set the intercepts for the two models.
```{r}
a_0 = 0
b_0 = 0
```

For pedagogical purposes, I will demonstrate the analysis on a single simulated dataset, then run the full MC study invisibly and only show the results. First, we choose a sample size, $n$, and generate $X$ and $W$, then use them to generate $M$ and $Y$.

```{r}
n = all_Ns[1]

X = rnorm(n, mean=0, sd=1)
W = matrix(rnorm(n*p_conf, mean=0, sd=1), nrow = n, ncol = p_conf)

# Generate M
e_M = rnorm(n, 0, sigma_M)
M = a_0 + a_1*X + W%*%A_2 + e_M

# Generate Y
e_Y = rnorm(n, 0, sigma_Y)
Y = b_0 + b_1*M + b_2*X + W%*%B_3 + e_Y
```

## Estimate Mediation Effect

Next, we fit regression models for $M$ and $Y$ and extract relevant output (in the loop version, these are stored at each iteration)

```{r}
M_data = data.frame(M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
M_model = lm(M ~ X + W1 + W2 + W3, data = M_data)

a_hat = summary(M_model)$coefficients[,1]
a_SE = summary(M_model)$coefficients[,2]
a_cov = vcov(M_model)

Y_data = data.frame(Y, M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
Y_model = lm(Y ~ M + X + W1 + W2 + W3, data = Y_data)

b_hat = summary(Y_model)$coefficients[,1]
b_SE = summary(Y_model)$coefficients[,2]
b_cov = vcov(Y_model)
```

Now we can extract relevant coefficients and estimate the mediation effect

```{r}
a_x = a_hat[2]
b_x = b_hat[3]
b_m = b_hat[2]

med_hat = a_x*b_m + b_x
```

## Estimate Standard Error

Finally, we can estimate the SE of the mediation effect using the $\delta$-method. The way I do this is excessive here since the mediation effect has such a simple formula. Later though, the extra structure will make our lives easier.

We start by constructing the joint covariance matrix of the regression coefficients from the two models. This matrix is block-diagonal, with the blocks corresponding to the covariance matrices of the coefficients from the two models and off-diagonal entries zero.

```{r}
a_length = nrow(a_cov)
b_length = nrow(b_cov)

joint_cov = matrix(0, nrow = a_length + b_length, ncol = a_length + b_length)
joint_cov[1:a_length, 1:a_length] = a_cov
joint_cov[(a_length+1):(a_length+b_length), (a_length+1):(a_length+b_length)] = b_cov
```

Next, we compute the gradient of the mediation effect with respect to each regression coefficient.

```{r}
grad_a_0 = 0
grad_a_1 = b_m
grad_A_2 = rep(0, times = p_conf)

grad_b_0 = 0
grad_b_1 = a_x
grad_b_2 = 1
grad_B_3 = rep(0, times = p_conf)

grad_med = c(grad_a_0, grad_a_1, grad_A_2, grad_b_0, grad_b_1, grad_b_2, grad_B_3)
```

Finally, we can use the delta method to estimate the SE of the mediation effect. Note that the $\delta$-method works on asymptotic covariances, so we need to make sure to multiply/divide by $n$ or $\sqrt{n}$ where appropriate.

```{r}
asymp_reg_cov = n * joint_cov

med_asymp_var = grad_med %*% asymp_reg_cov %*% grad_med
med_asymp_SE = sqrt(med_asymp_var)

med_SE = med_asymp_SE / sqrt(n)
```

Putting everything together, we have an estimated mediation effect of `r med_hat` with an estimated SE of `r med_SE`.

## Monte Carlo Study

We now repeat the above analysis `r num_reps` times for each of various values of $n$. We also re-run the boilerplate code (not shown)

```{r, include=FALSE}
<<boilerplate>>
```

```{r, cache=TRUE, dependson=c("boilerplate")}

for(j in seq_along(all_Ns)){
  n = all_Ns[j]

  # Containers for output with this value of n
  some_a_hats = list()
  some_b_hats = list()

  some_a_SEs = list()
  some_b_SEs = list()

  some_med_hats = list()
  some_med_SEs = list()

  for(i in 1:num_reps){
    # Generate data
    X = rnorm(n, mean=0, sd=1)
    W = matrix(rnorm(n*p_conf, mean=0, sd=1), nrow = n, ncol = p_conf)

    ## Generate M
    e_M = rnorm(n, 0, sigma_M)
    M = a_0 + a_1*X + W%*%A_2 + e_M

    ## Generate Y
    e_Y = rnorm(n, 0, sigma_Y)
    Y = b_0 + b_1*M + b_2*X + W%*%B_3 + e_Y

    # Fit models

    ## M
    M_data = data.frame(M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
    M_model = lm(M ~ X + W1 + W2 + W3, data = M_data)

    a_hat = summary(M_model)$coefficients[,1]
    a_SE = summary(M_model)$coefficients[,2]
    a_cov = vcov(M_model)

    some_a_hats[[i]] = a_hat
    some_a_SEs[[i]] = a_SE

    ## Y
    Y_data = data.frame(Y, M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
    Y_model = lm(Y ~ M + X + W1 + W2 + W3, data = Y_data)

    b_hat = summary(Y_model)$coefficients[,1]
    b_SE = summary(Y_model)$coefficients[,2]
    b_cov = vcov(Y_model)

    some_b_hats[[i]] = b_hat
    some_b_SEs[[i]] = b_SE

    # Estimate mediation effect

    ## Extract coefficients
    a_x = a_hat[2]
    b_x = b_hat[3]
    b_m = b_hat[2]

    ## Compute estimate
    med_hat = a_x*b_m + b_x
    some_med_hats[[i]] = med_hat

    # Estimate SE

    ## Build joint covariance matrix
    a_length = nrow(a_cov)
    b_length = nrow(b_cov)

    joint_cov = matrix(0, nrow = a_length + b_length, ncol = a_length + b_length)
    joint_cov[1:a_length, 1:a_length] = a_cov
    joint_cov[(a_length+1):(a_length+b_length), (a_length+1):(a_length+b_length)] = b_cov

    ## Compute gradient of mediation effect
    grad_a_0 = 0
    grad_a_1 = b_m
    grad_A_2 = rep(0, times = p_conf)

    grad_b_0 = 0
    grad_b_1 = a_x
    grad_b_2 = 1
    grad_B_3 = rep(0, times = p_conf)

    grad_med = c(grad_a_0, grad_a_1, grad_A_2, grad_b_0, grad_b_1, grad_b_2, grad_B_3)

    ## Apply delta-method
    asymp_reg_cov = n * joint_cov

    med_asymp_var = grad_med %*% asymp_reg_cov %*% grad_med
    med_asymp_SE = sqrt(med_asymp_var)

    med_SE = med_asymp_SE / sqrt(n)
    some_med_SEs[[i]] = med_SE
  }

  # Store output for current value of n
  all_a_hats[[j]] = some_a_hats
  all_b_hats[[j]] = some_b_hats

  all_a_SEs[[j]] = some_a_SEs
  all_b_SEs[[j]] = some_b_SEs

  all_med_hats[[j]] = some_med_hats
  all_med_SEs[[j]] = some_med_SEs
}

```

We now process the output from our simulation and summarize the results in a table. We give the mean and median of the estimated standard errors, as well as their relative errors (in $\%$) as estimates of the empirical standard error. See Table \ref{tab:results_cont_cont_fix}. Note that the relative error decreased and became more consistent when I increased the number of Monte Carlo replicates from 100 to 1000.

```{r}
# Re-format output to be more useful. 
data_med_hats = lapply(all_med_hats, unlist)
data_med_SEs = lapply(all_med_SEs, unlist)

# SD of estimates across Monte Carlo samples
emp_SEs = sapply(data_med_hats, sd)

# Summaries of estimated SEs (mean and median)
mean_delta_SEs = sapply(data_med_SEs, mean)
median_delta_SEs = sapply(data_med_SEs, median)

# Combine into a table
results = data.frame(n = all_Ns, Empirical = emp_SEs, 
                     Mean = mean_delta_SEs, Median = median_delta_SEs,
                     "Percent_Err_Mean" = 100*(mean_delta_SEs - emp_SEs)/emp_SEs,
                     "Percent_Err_Median" = 100*(median_delta_SEs - emp_SEs)/emp_SEs)

```

```{r results="asis", echo=FALSE}
## Print table
kbl(results, format = "latex", booktabs = T, caption = "Summary of SE estimates for mediation effect under continuous response, continuous mediator, fixed-effects model \\label{tab:results_cont_cont_fix}")
```

# Continuous Response, Binary Mediator, Fixed-Effects

```{r, include=FALSE}
# Run the boilerplate setup
<<boilerplate>>
```

We now generate $M$ as a binary variable using a logistic regression model. We continue to use a continuous model for $Y|M$. This analysis is sufficiently complicated that I wrote some helper functions in another script. See `Helpers.R`.

```{r}
source("../src/Helpers.R")
```

## Generate data

The continuous variable is modelled as linear predictor plus residual. We therefore need to set the residual variance.

```{r}
sigma_Y = 0.2
```

We also need to set the intercepts for the two models. Recall that we want the mean of the linear predictor to be zero in both models, and that we approximate the mean of $M$ as the logit of the mean of its linear predictor (i.e. $\mathbb{E}M \approx \mathrm{expit}(0) = 0.5$).
```{r}
a_0 = 0
b_0 = -0.5
```

As in the previous section, I demonstrate the analysis on a single simulated dataset, then run the full MC study invisibly and only show the results. First, we choose a sample size, $n$, and generate $X$ and $W$, then use them to generate $M$ and $Y$.

```{r}
n = all_Ns[1]

X = rnorm(n, mean=0, sd=1)
W = matrix(rnorm(n*p_conf, mean=0, sd=1), nrow = n, ncol = p_conf)

# Generate M
eta_vec = a_0 + a_1*X + W%*%A_2
p_M_vec = expit(eta_vec)
M = rbinom(n, size = 1, prob = p_M_vec)

# Generate Y
e_Y = rnorm(n, 0, sigma_Y)
Y = b_0 + b_1*M + b_2*X + W%*%B_3 + e_Y
```

## Estimate Mediation Effect

Next, we fit regression models for $M$ and $Y$ and extract relevant output (in the loop version, these are stored at each iteration)

```{r}
M_data = data.frame(M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
M_model = glm(M ~ X + W1 + W2 + W3, data = M_data, family = binomial(link = "logit"))

a_hat = summary(M_model)$coefficients[,1]
a_SE = summary(M_model)$coefficients[,2]
a_cov = vcov(M_model)

Y_data = data.frame(Y, M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
Y_model = lm(Y ~ M + X + W1 + W2 + W3, data = Y_data)

b_hat = summary(Y_model)$coefficients[,1]
b_SE = summary(Y_model)$coefficients[,2]
b_cov = vcov(Y_model)
```

Now we can extract relevant coefficients and estimate the mediation effect. Note that this is more involved than the fully continuous case. In particular, the total effect of $X$ on $Y$ depends on the levels of $X$ and the confounders, $W$. I evaluate the effect at $X=0$ and $W = [1,1,1]$. The former represents the effect of a binary exposure, but the latter is chosen completely arbitrarily.

```{r}
x_pred = 0
W_pred = c(1,1,1)

a_0_hat = a_hat[1]
a_x_hat = a_hat[2]
A_2_hat = a_hat[3:5]

b_0_hat = b_hat[1]
b_m_hat = b_hat[2]
b_x_hat = b_hat[3]
B_3_hat = b_hat[4:6]

# Linear predictor for M
eta_hat = a_0_hat + a_x_hat*x_pred + W_pred%*%A_2_hat

# Increment in the conditional expectation of M
delta_hat = get_delta(eta_hat, a_x_hat)

# Increment in the conditional expectation of Y
med_hat = get_gamma(delta_hat, b_m_hat, b_x_hat)
```

## Estimate Standard Error

Finally, we can estimate the SE of the mediation effect using the $\delta$-method. Now that we have a more complicated expression for the total effect, it makes more sense to use the general formula for the $\delta$-method standard error.

We start by constructing the joint covariance matrix of the regression coefficients from the two models. This matrix is block-diagonal, with the blocks corresponding to the covariance matrices of the coefficients from the two models and off-diagonal entries zero.

```{r}
a_length = nrow(a_cov)
b_length = nrow(b_cov)

joint_cov = matrix(0, nrow = a_length + b_length, ncol = a_length + b_length)
joint_cov[1:a_length, 1:a_length] = a_cov
joint_cov[(a_length+1):(a_length+b_length), (a_length+1):(a_length+b_length)] = b_cov
```

Next, we compute the gradient of the mediation effect with respect to each regression coefficient.

```{r}
d_gamma_d_theta(eta_hat, x_pred, W_pred, a_hat, b_hat)
```

Finally, we can use the delta method to estimate the SE of the mediation effect. Note that the $\delta$-method works on asymptotic covariances, so we need to make sure to multiply/divide by $n$ or $\sqrt{n}$ where appropriate.

```{r}
asymp_reg_cov = n * joint_cov

med_asymp_var = grad_med %*% asymp_reg_cov %*% grad_med
med_asymp_SE = sqrt(med_asymp_var)

med_SE = med_asymp_SE / sqrt(n)
```

Putting everything together, we have an estimated mediation effect of `r med_hat` with an estimated SE of `r med_SE`.

## Monte Carlo Study

We now repeat the above analysis `r num_reps` times for each of various values of $n$. Values of $X$ and $W$ for which we compute the total effect are set at the beginning. We also re-run the boilerplate code (not shown)

```{r, include=FALSE}
<<boilerplate>>
```

```{r}
# Values of X and W for which we compute the total effect
x_pred = 0
W_pred = c(1,1,1)
```

```{r, cache=TRUE, dependson=c("boilerplate")}

for(j in seq_along(all_Ns)){
  n = all_Ns[j]

  # Containers for output with this value of n
  some_a_hats = list()
  some_b_hats = list()

  some_a_SEs = list()
  some_b_SEs = list()

  some_med_hats = list()
  some_med_SEs = list()

  for(i in 1:num_reps){
    # Generate data
    X = rnorm(n, mean=0, sd=1)
    W = matrix(rnorm(n*p_conf, mean=0, sd=1), nrow = n, ncol = p_conf)

    ## Generate M
    eta_vec = a_0 + a_1*X + W%*%A_2
    p_M_vec = expit(eta_vec)
    M = rbinom(n, size = 1, prob = p_M_vec)

    ## Generate Y
    e_Y = rnorm(n, 0, sigma_Y)
    Y = b_0 + b_1*M + b_2*X + W%*%B_3 + e_Y

    # Fit models

    ## M
    M_data = data.frame(M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
    M_model = glm(M ~ X + W1 + W2 + W3, data = M_data, family = binomial(link = "logit"))

    a_hat = summary(M_model)$coefficients[,1]
    a_SE = summary(M_model)$coefficients[,2]
    a_cov = vcov(M_model)

    some_a_hats[[i]] = a_hat
    some_a_SEs[[i]] = a_SE

    ## Y
    Y_data = data.frame(Y, M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
    Y_model = lm(Y ~ M + X + W1 + W2 + W3, data = Y_data)

    b_hat = summary(Y_model)$coefficients[,1]
    b_SE = summary(Y_model)$coefficients[,2]
    b_cov = vcov(Y_model)

    some_b_hats[[i]] = b_hat
    some_b_SEs[[i]] = b_SE

    # Estimate mediation effect

    ## Extract coefficients
    a_0_hat = a_hat[1]
    a_x_hat = a_hat[2]
    A_2_hat = a_hat[3:length(a_hat)]

    b_0_hat = b_hat[1]
    b_m_hat = b_hat[2]
    b_x_hat = b_hat[3]
    B_3_hat = b_hat[4:length(b_hat)]

    ## Linear predictor for M
    eta_hat = a_0_hat + a_x_hat*x_pred + W_pred%*%A_2_hat

    ## Increment in the conditional expectation of M
    delta_hat = get_delta(eta_hat, a_x_hat)

    ## Increment in the conditional expectation of Y (i.e. mediation effect)
    med_hat = get_gamma(delta_hat, b_m_hat, b_x_hat)
    some_med_hats[[i]] = med_hat

    # Estimate SE

    ## Build joint covariance matrix
    a_length = nrow(a_cov)
    b_length = nrow(b_cov)

    joint_cov = matrix(0, nrow = a_length + b_length, ncol = a_length + b_length)
    joint_cov[1:a_length, 1:a_length] = a_cov
    joint_cov[(a_length+1):(a_length+b_length), (a_length+1):(a_length+b_length)] = b_cov

    ## Compute gradient of mediation effect
    grad_med = d_gamma_d_theta(eta_hat, x_pred, W_pred, a_hat, b_hat)

    ## Apply delta-method
    asymp_reg_cov = n * joint_cov

    med_asymp_var = grad_med %*% asymp_reg_cov %*% grad_med
    med_asymp_SE = sqrt(med_asymp_var)

    med_SE = med_asymp_SE / sqrt(n)
    some_med_SEs[[i]] = med_SE
  }

  # Store output for current value of n
  all_a_hats[[j]] = some_a_hats
  all_b_hats[[j]] = some_b_hats

  all_a_SEs[[j]] = some_a_SEs
  all_b_SEs[[j]] = some_b_SEs

  all_med_hats[[j]] = some_med_hats
  all_med_SEs[[j]] = some_med_SEs
}

```

We now process the output from our simulation and summarize the results in a table. We give the mean and median of the estimated standard errors, as well as their relative errors (in $\%$) as estimates of the empirical standard error. See Table \ref{tab:results_cont_bin_fix}. Note that the relative error decreased and became more consistent when I increased the number of Monte Carlo replicates from 100 to 1000.

```{r}
# Re-format output to be more useful. 
data_med_hats = lapply(all_med_hats, unlist)
data_med_SEs = lapply(all_med_SEs, unlist)

# SD of estimates across Monte Carlo samples
emp_SEs = sapply(data_med_hats, sd)

# Summaries of estimated SEs (mean and median)
mean_delta_SEs = sapply(data_med_SEs, mean)
median_delta_SEs = sapply(data_med_SEs, median)

# Combine into a table
results = data.frame(n = all_Ns, Empirical = emp_SEs, 
                     Mean = mean_delta_SEs, Median = median_delta_SEs,
                     "Percent_Err_Mean" = 100*(mean_delta_SEs - emp_SEs)/emp_SEs,
                     "Percent_Err_Median" = 100*(median_delta_SEs - emp_SEs)/emp_SEs)

```

```{r results="asis", echo=FALSE}
## Print table
kbl(results, format = "latex", booktabs = T, caption = "Summary of SE estimates for mediation effect under continuous response, continuous mediator, fixed-effects model \\label{tab:results_cont_bin_fix}")
```